### 2022-03-23 race compas analysis.Rmd
knitting rmd

---
title: "COMPAS Race Analysis"
author: "Elisabeth Nesmith"
date: "3/8/2022"
output: html_document
---

```{r setup, include=FALSE}
library(reticulate)
```

```{python}
import matplotlib.pyplot as plt
import matplotlib.patches as patches
import numpy as np
import pandas as pd
import seaborn as sns

import matplotlib.pyplot as plt
import matplotlib.patches as mpatches

from sklearn.compose import make_column_transformer
from sklearn.linear_model import LogisticRegression
from sklearn.linear_model import LogisticRegressionCV

from sklearn.model_selection import GridSearchCV, train_test_split
from sklearn.preprocessing import OneHotEncoder

from aif360.datasets import CompasDataset
from aif360.datasets import BinaryLabelDataset
from aif360.sklearn.datasets import fetch_compas
from aif360.algorithms.inprocessing import MetaFairClassifier

from aif360.sklearn.preprocessing import ReweighingMeta
from aif360.sklearn.inprocessing import AdversarialDebiasing
from aif360.sklearn.postprocessing import CalibratedEqualizedOdds, PostProcessingMeta
from aif360.sklearn.metrics import disparate_impact_ratio, average_odds_error, generalized_fpr, average_odds_difference
from aif360.sklearn.metrics import generalized_fnr, difference, statistical_parity_difference, equal_opportunity_difference, generalized_entropy_error

from IPython.display import Markdown, display
```

```{python}
## load in data
## In machine learning tasks, specifically with supervised learning, you have features and labels. 
## The features are the descriptive attributes (they are defined as X), and the label (y) is what you're attempting to predict or forecast

X, y = fetch_compas()
print(f'There are {X.shape[0]} entries and {X.shape[1]} features')
X.head()
```

```{python}
## because our analysis is mainly focusing on how the algorithm treats white and Black people differently, we are
## dropping the rows of data where race != Caucasian or African American
X_new = X[(X.race == "Caucasian") | (X.race == "African-American")]
print(f'There are {X_new.shape[0]} entries and {X_new.shape[1]} features')
X_new.head()
```

```{python}
## drop unused race categories
# list of categories to be removed
X_new["race"] = X_new["race"].cat.remove_unused_categories()
```

```{python}
y_new = y[(y.index.get_level_values(2) == "Caucasian") | (y.index.get_level_values(2) == "African-American")]
y_new.head()
```

```{python}
# Function for visualising the confusion matrix and other statistics
# https://github.com/DTrimarchi10/confusion_matrix/blob/master/cf_matrix.py

def make_confusion_matrix(cf_matrix, model):
  group_names = ["True Negative","False Positive","False Negative","True Positive"]
  group_counts = ["{0:0.0f}".format(value) for value in
                  cf_matrix.flatten()]
  group_percentages = ["{0:.2%}".format(value) for value in
                      cf_matrix.flatten()/np.sum(cf_matrix)]

  group_labels = ["{}\n".format(value) for value in group_names]
  group_counts = ["{0:0.0f}\n".format(value) for value in cf_matrix.flatten()]
  group_percentages = ["{0:.2%}".format(value) for value in cf_matrix.flatten()/np.sum(cf_matrix)]

  box_labels = [f"{v1}{v2}{v3}".strip() for v1, v2, v3 in zip(group_labels,group_counts,group_percentages)]
  box_labels = np.asarray(box_labels).reshape(cf_matrix.shape[0],cf_matrix.shape[1])

  # add more statistics
  accuracy  = np.trace(cf_matrix) / float(np.sum(cf_matrix))
  precision = cf_matrix[1,1] / sum(cf_matrix[:,1])
  recall    = cf_matrix[1,1] / sum(cf_matrix[1,:])
  f1_score  = 2*precision*recall / (precision + recall)
  stats_text = "\n\nAccuracy={:0.3f}\nPrecision={:0.3f}\nRecall={:0.3f}\nF1 Score={:0.3f}".format(
      accuracy,precision,recall,f1_score)

  categories=["Survived", "Recidivated"]
  sns.heatmap(cf_matrix,annot=box_labels,fmt="",cmap='Purples',xticklabels=categories,yticklabels=categories)

  plt.ylabel('True label')
  plt.xlabel('Predicted label' + stats_text)
  plt.title(f"Confusion matrix and statistics for the {model} model");

## defining function for displaying metrics of training and test data by sex 
def metrics_per_group(y_test, y_pred):
	# y true per group
	y_test_white = y_test.loc[y_test.index.get_level_values(2) == 1]
	y_test_black = y_test.loc[y_test.index.get_level_values(2) == 0]

	# y_pred per group
	y_pred_white = y_pred[y_test.index.get_level_values(2) == 1]
	y_pred_black = y_pred[y_test.index.get_level_values(2) == 0]

	# metrics
	scores = []
	scores.append(accuracy_score(y_test, y_pred))
	scores.append(recall_score(y_test, y_pred))
	scores.append(precision_score(y_test, y_pred))

	scores.append(accuracy_score(y_test_black, y_pred_black))
	scores.append(recall_score(y_test_black, y_pred_black))
	scores.append(precision_score(y_test_black, y_pred_black))

	scores.append(accuracy_score(y_test_white, y_pred_white))
	scores.append(recall_score(y_test_white, y_pred_white))
	scores.append(precision_score(y_test_white, y_pred_white))

	attribute = ["all"]*3 + ["black"] *3 + ["white"] *3
	metric = ["accuracy", "recall", "precision"] * 3

	# dictionary of lists 
	dict = {'race': attribute, 'metrics': metric, 'score': scores}

	df = pd.DataFrame(dict)

	sns.barplot(x = "metrics", y = "score", hue = "race", data = df, palette = ['#dfcd1a', '#9d0677', '#236c48'])
	plt.title("Performance metrics by groups")

def plot_fair_metrics(fair_metrics_mitigated, model): 
  cols = ['statistical_parity_difference', 'equal_opportunity_difference', 'generalized_entropy']
  obj_fairness = [[0,0,1]]

  # row for objectives    
  fair_metrics = pd.DataFrame(data=obj_fairness, index=['objective'], columns=cols)

  # row for baseline model
  fair_metrics.loc['Baseline Model'] = [stat_par_diff, eq_opp_diff, gen_entr_error]

  # row for mitigated bias
  fair_metrics.loc[model] = fair_metrics_mitigated

  metrics_len = len(cols)

  fig, ax = plt.subplots(figsize=(20,4), ncols=metrics_len, nrows=1)

  plt.subplots_adjust(
      left    =  0.125, 
      bottom  =  0.1, 
      right   =  0.9, 
      top     =  0.9, 
      wspace  =  .5, 
      hspace  =  1.1
  )

  y_title_margin = 1.2

  plt.suptitle("Fairness metrics", y = 1.09, fontsize=20)
  sns.set(style="dark")

  cols = fair_metrics.columns.values
  obj = fair_metrics.loc['objective']
  size_rect = [0.2,0.2,0.2,0.4,0.25]
  rect = [-0.1,-0.1,-0.1,0.8,0]
  bottom = [-1,-1,-1,0,0]
  top = [1,1,1,2,1]
  bound = [[-0.1,0.1],[-0.1,0.1],[-0.1,0.1],[0.8,1.2],[0,0.25]]

  for i in range(0,metrics_len):
      plt.subplot(1, metrics_len, i+1)
      ax = sns.barplot(x=fair_metrics.index[1:len(fair_metrics)], y=fair_metrics.iloc[1:len(fair_metrics)][cols[i]])

      for j in range(0,len(fair_metrics)-1):
          a, val = ax.patches[j], fair_metrics.iloc[j+1][cols[i]]
          marg = -0.2 if val < 0 else 0.1
          ax.text(a.get_x()+a.get_width()/5, a.get_y()+a.get_height()+marg, round(val, 3), fontsize=15,color='black')

      plt.ylim(bottom[i], top[i])
      plt.setp(ax.patches, linewidth=0)
      ax.add_patch(patches.Rectangle((-5,rect[i]), 10, size_rect[i], alpha=0.3, facecolor="green", linewidth=1, linestyle='solid'))
      plt.axhline(obj[i], color='black', alpha=0.3)
      plt.title(cols[i])
      ax.set_ylabel('')    
      ax.set_xlabel('')
```

```{python}
X_new.index = pd.MultiIndex.from_arrays(X_new.index.codes, names=X_new.index.names)
y_new.index = pd.MultiIndex.from_arrays(y_new.index.codes, names=y_new.index.names)
# 0 is African American, 2 is Caucasian
```

```{python}
# set caucasian equal to 1 instead of 2
X_new = X_new.rename(index={2: 1}, level='race')
X_new
```

```{python}
# set target class to 0/1 
y_new = pd.Series(y_new.factorize(sort=True)[0], index=y_new.index)
# set caucasian equal to 1 instead of 2
y_new = y_new.rename(index={2: 1}, level='race')
```

```{python}
y_new
```

# Exploratory Data Analysis
```{python}
df_viz = X_new.copy()
df_viz['race'] = X_new['race'].replace({1.0: 'Caucasian', 0.0: 'African-American'})
df_viz['two_year_recid'] = y_new.replace({1:'Recidivated', 0: 'Survived'})
df_viz.index = df_viz.index.droplevel('race')

purple = '#9d0677'
green = '#30875c'
workshop_palette = [purple, green]
df_viz.head()
```

```{python}
# barplot of recividism
sns.countplot(x='two_year_recid', data=df_viz, palette=workshop_palette)
plt.title('Two Year Recidivism Rate')
```

```{python}
# barplot of race
sns.countplot(y="race", data=df_viz, palette=["#30875c"])
plt.title('Race Distribution (White and Black individuals only)')
```

```{python}
# age distribution by race
ax = sns.kdeplot(x="age", hue="race", data=df_viz);
kdeline = ax.lines[0]
mean_fem = df_viz.groupby('race').age.median()[0]
xs = kdeline.get_xdata()
ys = kdeline.get_ydata()
height_fem = np.interp(mean_fem, xs, ys)
ax.vlines(mean_fem, 0, height_fem, color=purple, ls=':')

mean_m = df_viz.groupby('race').age.median()[1]
height_m = np.interp(mean_m, xs, ys)
ax.vlines(mean_m, 0, height_m, color=green, ls=':')

print(df_viz.groupby('race').age.median())

plt.title('Distribution of Age by Race')
```

```{python}
# recividism by race
by_sex = sns.countplot(x="race", hue="two_year_recid", data=df_viz, palette=workshop_palette)

plt.title('Two Year Recidivism by Race')
plt.show()
```

# Baseline Metrics

```{python}
X_train, X_test, y_train, y_test = train_test_split(X_new, y_new, random_state=1234567)

# one-hoy encode the categorical features
data_preproc = make_column_transformer(
        (OneHotEncoder(sparse=False, handle_unknown='ignore'), X_train.dtypes == 'category'))

X_train = pd.DataFrame(data_preproc.fit_transform(X_train), index=X_train.index)
X_test = pd.DataFrame(data_preproc.transform(X_test), index=X_test.index)

# to save the information for the column names
pd.get_dummies(X_new).head()
```

```{python}
from aif360.sklearn.preprocessing import ReweighingMeta, Reweighing
from sklearn.metrics import accuracy_score, confusion_matrix, plot_roc_curve, recall_score, precision_score

lr = LogisticRegressionCV(solver='lbfgs')
reg = lr.fit(X_train, y_train)
y_pred = reg.predict(X_test)

acc_base = accuracy_score(y_test, y_pred)
print(f'[Baseline] The test accuracy of the algorithm is: {acc_base: .2%}')
```

```{python}
##Look at 2nd column of matrix for precision 
##2nd row of matrix for recall 
cf_matrix = confusion_matrix(y_test, y_pred)
make_confusion_matrix(cf_matrix, "[Baseline]")
plt.show()
```

```{python}
metrics_per_group(y_test, y_pred)
plt.show()
```

In the Black population, the model correctly identifies 70% of people who recidivated. In the white population, the model correctly identifies 20% of people who recidivated.

### (Group fairness) Statistical Parity Difference

Computed as the difference of the rate of favorable outcomes (in this case, did not recidivate) received by the unprivileged group to the privileged group. It essentially equalizes the outcomes across the protected and non-protected groups.
The ideal value of this metric is 0. Fairness for this metric is between -0.1 and 0.1. A negative value means there is higher benefit for the privileged group (in this case, white defendants).

```{python}
stat_par_diff = statistical_parity_difference(y_test, prot_attr='race', pos_label = 0)

print(f'[Baseline] The statistical parity difference is {stat_par_diff: .2}')
```

### Generalized entropy error

Generalized entropy index is proposed as a unified individual and group fairness measure. It measures the inequality in benefit allocation for individuals.
A value of 0 implies perfect fairness.
Fairness is indicated by lower scores, higher scores are problematic.

```{python}
gen_entr_error = generalized_entropy_error(y_test, y_pred, alpha=1, pos_label = 0)

print(f'[Baseline] The generalized entropy error is {gen_entr_error: .2}')
```

### (Group fairness) Equal Opportunity Difference

This metric is computed as the difference of true positive rates between the unprivileged and the privileged groups. The true positive rate is the ratio of true positives to the total number of actual positives for a given group.
The ideal value is 0. A value of < 0 implies higher benefit for the privileged group and a value > 0 implies higher benefit for the unprivileged group.
Fairness for this metric is between -0.1 and 0.1

```{python}
eq_opp_diff = equal_opportunity_difference(y_test, y_pred, prot_attr='race', pos_label = 0)

print(f'[Baseline] The equal opportunity difference is {eq_opp_diff: .2}')
```

```{python}
plot_fair_metrics([0, 0, 0], '')
plt.show()
```

# Pre-Processing Approach
```{python}
lr = LogisticRegressionCV(solver='lbfgs')
rew = ReweighingMeta(estimator=lr, reweigher=Reweighing('race'))
rew.fit(X_train, y_train)
y_pred_REW = rew.predict(X_test)
```

```{python}
acc_REW = accuracy_score(y_test, y_pred_REW)
print(f'[Reweighting] The test accuracy of the algorithm is: {acc_REW: .2%}')
```

```{python}
cf_matrix = confusion_matrix(y_test, y_pred_REW)
make_confusion_matrix(cf_matrix, "[Reweighting]")
plt.show()
```

```{python}
metrics_per_group(y_test, y_pred_REW)
plt.show()
```

```{python}
stat_par_diff_RW = statistical_parity_difference(y_test, y_pred_REW, prot_attr='race', pos_label = 0)

print(f'[Reweighting] The statistical parity difference is {stat_par_diff_RW: .2}')
```

This is a large improvement over our baseline model, which was -0.14. It still implies a slight benefit for white individuals, but it is in the range of -0.1 and 0.1.

```{python}
eq_opp_diff_RW = equal_opportunity_difference(y_test, y_pred_REW, prot_attr='race', pos_label = 0)

print(f'[Reweighting] The equal opportunity difference is {eq_opp_diff_RW: .2}')
```

This is also an improvement from the baseline, as it is now between -0.1 and 0.1. It also now implies a slight benefit for Black defendants.

```{python}
gen_entr_error_RW = generalized_entropy_error(y_test, y_pred_REW, alpha=1, pos_label = 0)

print(f'[Reweighting] The generalized entropy error is {gen_entr_error_RW: .2}')
```

This is only a very slight improvement from 0.24 to 0.23.

```{python}
plot_fair_metrics([stat_par_diff_RW, eq_opp_diff_RW, gen_entr_error_RW], 'Reweighting')
plt.show()
```

# In-Processing

```{python}
df_train = X_train.copy()
df_train["two_year_recid"] = y_train
df_train.rename(columns={6:'race'}, inplace=True)
df_train = df_train.reset_index(drop=True)

df_test = X_test.copy()
df_test["two_year_recid"] = y_test
df_test = df_test.reset_index(drop=True)
df_test.rename(columns={6:'race'}, inplace=True)
```

```{python}
train_BLD = BinaryLabelDataset(favorable_label='0',
                                unfavorable_label='1',
                                df=df_train,
                                label_names=['two_year_recid'],
                                protected_attribute_names=['race'])

test_BLD = BinaryLabelDataset(favorable_label='0',
                                unfavorable_label='1',
                                df=df_test,
                                label_names=['two_year_recid'],
                                protected_attribute_names=['race'])
```

```{python}
# meta = MetaFairClassifier(tau=0.8, sensitive_attr='race', type='fdr', seed=1234567)
# 
# meta.fit(train_BLD)
# y_pred_META = meta.predict(test_BLD)
# y_pred_META = y_pred_META.labels.flatten()
```

```{python}
# acc_AD = accuracy_score(y_test, y_pred_META)
# print(f'[ Meta-Classification] The test accuracy of the algorithm is: {acc_AD: .2%}')
```

# Post-Processing

```{python}
from aif360.sklearn.postprocessing import CalibratedEqualizedOdds, PostProcessingMeta

pp = CalibratedEqualizedOdds('race', cost_constraint='fnr', random_state=1234567)
ceo = PostProcessingMeta(estimator=lr, postprocessor=pp, random_state=1234567)
ceo.fit(X_train, y_train)
y_pred_CEO = ceo.predict(X_test)
y_proba_CEO = ceo.predict_proba(X_test)
```

```{python}
acc_CEO = accuracy_score(y_test, y_pred_CEO)
print(f'[Calibrated Equalized Odds] The test accuracy of the algorithm is: {acc_CEO: .2%}')
```

```{python}
cf_matrix = confusion_matrix(y_test, y_pred_CEO)
make_confusion_matrix(cf_matrix, "[Calibrated Equalized Odds]")
plt.show()
```

```{python}
stat_par_diff_CEO = statistical_parity_difference(y_test, y_pred_CEO, prot_attr='race', pos_label = 0)

print(f'[Calibrated Equalized Odds] The statistical parity difference is {stat_par_diff_CEO: .2}')
```

```{python}
eq_opp_diff_CEO = equal_opportunity_difference(y_test, y_pred_CEO, prot_attr='race', pos_label = 0)

print(f'[Calibrated Equalized Odds] The equal opportunity difference is {eq_opp_diff_CEO: .2}')
```

```{python}
gen_entr_error_CEO = generalized_entropy_error(y_test, y_pred_CEO, alpha=1, pos_label = 0)

print(f'[Calibrated Equalized Odds] The generalized entropy error is {gen_entr_error_CEO: .2}')
```

```{python}
plot_fair_metrics([stat_par_diff_CEO, eq_opp_diff_CEO, gen_entr_error_CEO], 'Calibrated Equalized Odds')
plt.show()
```