### 2022-04-05 race compas analysis fairness metrics.Rmd
label chunks

---
title: "COMPAS Race Analysis and Fairness Metrics"
author: "Elisabeth Nesmith"
date: "3/8/2022"
output: html_document
---

# Set Up

```{r py setup, include=FALSE}
library(reticulate)
```

```{python import, include= FALSE}
import matplotlib.pyplot as plt
import matplotlib.patches as patches
import numpy as np
import pandas as pd
import seaborn as sns

import matplotlib.pyplot as plt
import matplotlib.patches as mpatches

from sklearn.compose import make_column_transformer
from sklearn.linear_model import LogisticRegression
from sklearn.linear_model import LogisticRegressionCV

from sklearn.model_selection import GridSearchCV, train_test_split
from sklearn.preprocessing import OneHotEncoder

from aif360.datasets import CompasDataset
from aif360.datasets import BinaryLabelDataset
from aif360.sklearn.datasets import fetch_compas
from aif360.algorithms.inprocessing import MetaFairClassifier

from aif360.sklearn.preprocessing import ReweighingMeta
from aif360.sklearn.inprocessing import AdversarialDebiasing
from aif360.sklearn.postprocessing import CalibratedEqualizedOdds, PostProcessingMeta
from aif360.sklearn.metrics import disparate_impact_ratio, average_odds_error, generalized_fpr, average_odds_difference
from aif360.sklearn.metrics import generalized_fnr, difference, statistical_parity_difference, equal_opportunity_difference, generalized_entropy_error

from IPython.display import Markdown, display
```

# Data

```{python fetch compas}
## load in data
## In machine learning tasks, specifically with supervised learning, you have features and labels. 
## The features are the descriptive attributes (they are defined as X), and the label (y) is what you're attempting to predict or forecast

X, y = fetch_compas()
print(f'There are {X.shape[0]} entries and {X.shape[1]} features')
X.head()
```

## Filter/configure race

```{python filter race}
## because our analysis is mainly focusing on how the algorithm treats white and Black people differently, we are
## dropping the rows of data where race != Caucasian or African American
X_new = X[(X.race == "Caucasian") | (X.race == "African-American")]
print(f'There are {X_new.shape[0]} entries and {X_new.shape[1]} features')
X_new.head()
```

```{python drop other races}
## drop unused race categories
# list of categories to be removed
X_new["race"] = X_new["race"].cat.remove_unused_categories()
```

```{python configure races}
y_new = y[(y.index.get_level_values(2) == "Caucasian") | (y.index.get_level_values(2) == "African-American")]
y_new.head()
```

```{python initial confusion matrix}
# Function for visualising the confusion matrix and other statistics
# https://github.com/DTrimarchi10/confusion_matrix/blob/master/cf_matrix.py

def make_confusion_matrix(cf_matrix, model):
  group_names = ["True Negative","False Positive","False Negative","True Positive"]
  group_counts = ["{0:0.0f}".format(value) for value in
                  cf_matrix.flatten()]
  group_percentages = ["{0:.2%}".format(value) for value in
                      cf_matrix.flatten()/np.sum(cf_matrix)]

  group_labels = ["{}\n".format(value) for value in group_names]
  group_counts = ["{0:0.0f}\n".format(value) for value in cf_matrix.flatten()]
  group_percentages = ["{0:.2%}".format(value) for value in cf_matrix.flatten()/np.sum(cf_matrix)]

  box_labels = [f"{v1}{v2}{v3}".strip() for v1, v2, v3 in zip(group_labels,group_counts,group_percentages)]
  box_labels = np.asarray(box_labels).reshape(cf_matrix.shape[0],cf_matrix.shape[1])

  # add more statistics
  accuracy  = np.trace(cf_matrix) / float(np.sum(cf_matrix))
  precision = cf_matrix[1,1] / sum(cf_matrix[:,1])
  recall    = cf_matrix[1,1] / sum(cf_matrix[1,:])
  f1_score  = 2*precision*recall / (precision + recall)
  stats_text = "\n\nAccuracy={:0.3f}\nPrecision={:0.3f}\nRecall={:0.3f}\nF1 Score={:0.3f}".format(
      accuracy,precision,recall,f1_score)

  categories=["Survived", "Recidivated"]
  sns.heatmap(cf_matrix,annot=box_labels,fmt="",cmap='Purples',xticklabels=categories,yticklabels=categories)

  plt.ylabel('True label')
  plt.xlabel('Predicted label' + stats_text)
  plt.title(f"Confusion matrix and statistics for the {model} model");

## defining function for displaying metrics of training and test data by sex 
def metrics_per_group(y_test, y_pred):
	# y true per group
	y_test_white = y_test.loc[y_test.index.get_level_values(2) == 1]
	y_test_black = y_test.loc[y_test.index.get_level_values(2) == 0]

	# y_pred per group
	y_pred_white = y_pred[y_test.index.get_level_values(2) == 1]
	y_pred_black = y_pred[y_test.index.get_level_values(2) == 0]

	# metrics
	scores = []
	scores.append(accuracy_score(y_test, y_pred))
	scores.append(recall_score(y_test, y_pred))
	scores.append(precision_score(y_test, y_pred))

	scores.append(accuracy_score(y_test_black, y_pred_black))
	scores.append(recall_score(y_test_black, y_pred_black))
	scores.append(precision_score(y_test_black, y_pred_black))

	scores.append(accuracy_score(y_test_white, y_pred_white))
	scores.append(recall_score(y_test_white, y_pred_white))
	scores.append(precision_score(y_test_white, y_pred_white))

	attribute = ["all"]*3 + ["black"] *3 + ["white"] *3
	metric = ["accuracy", "recall", "precision"] * 3

	# dictionary of lists 
	dict = {'race': attribute, 'metrics': metric, 'score': scores}

	df = pd.DataFrame(dict)

	sns.barplot(x = "metrics", y = "score", hue = "race", data = df, palette = ['#dfcd1a', '#9d0677', '#236c48'])
	plt.title("Performance metrics by groups")

def plot_fair_metrics(fair_metrics_mitigated, model): 
  cols = ['statistical_parity_difference', 'equal_opportunity_difference', 'generalized_entropy']
  obj_fairness = [[0,0,1]]

  # row for objectives    
  fair_metrics = pd.DataFrame(data=obj_fairness, index=['objective'], columns=cols)

  # row for baseline model
  fair_metrics.loc['Baseline Model'] = [stat_par_diff, eq_opp_diff, gen_entr_error]

  # row for mitigated bias
  fair_metrics.loc[model] = fair_metrics_mitigated

  metrics_len = len(cols)

  fig, ax = plt.subplots(figsize=(20,4), ncols=metrics_len, nrows=1)

  plt.subplots_adjust(
      left    =  0.125, 
      bottom  =  0.1, 
      right   =  0.9, 
      top     =  0.9, 
      wspace  =  .5, 
      hspace  =  1.1
  )

  y_title_margin = 1.2

  plt.suptitle("Fairness metrics", y = 1.09, fontsize=20)
  sns.set(style="dark")

  cols = fair_metrics.columns.values
  obj = fair_metrics.loc['objective']
  size_rect = [0.2,0.2,0.2,0.4,0.25]
  rect = [-0.1,-0.1,-0.1,0.8,0]
  bottom = [-1,-1,-1,0,0]
  top = [1,1,1,2,1]
  bound = [[-0.1,0.1],[-0.1,0.1],[-0.1,0.1],[0.8,1.2],[0,0.25]]

  for i in range(0,metrics_len):
      plt.subplot(1, metrics_len, i+1)
      ax = sns.barplot(x=fair_metrics.index[1:len(fair_metrics)], y=fair_metrics.iloc[1:len(fair_metrics)][cols[i]])

      for j in range(0,len(fair_metrics)-1):
          a, val = ax.patches[j], fair_metrics.iloc[j+1][cols[i]]
          marg = -0.2 if val < 0 else 0.1
          ax.text(a.get_x()+a.get_width()/5, a.get_y()+a.get_height()+marg, round(val, 3), fontsize=15,color='black')

      plt.ylim(bottom[i], top[i])
      plt.setp(ax.patches, linewidth=0)
      ax.add_patch(patches.Rectangle((-5,rect[i]), 10, size_rect[i], alpha=0.3, facecolor="green", linewidth=1, linestyle='solid'))
      plt.axhline(obj[i], color='black', alpha=0.3)
      plt.title(cols[i])
      ax.set_ylabel('')    
      ax.set_xlabel('')
```

```{python new race index}
X_new.index = pd.MultiIndex.from_arrays(X_new.index.codes, names=X_new.index.names)
y_new.index = pd.MultiIndex.from_arrays(y_new.index.codes, names=y_new.index.names)
# 0 is African American, 2 is Caucasian
```

```{python re-index race}
# set caucasian equal to 1 instead of 2
X_new = X_new.rename(index={2: 1}, level='race')
X_new
```

```{python target class}
# set target class to 0/1 
y_new = pd.Series(y_new.factorize(sort=True)[0], index=y_new.index)
# set caucasian equal to 1 instead of 2
y_new = y_new.rename(index={2: 1}, level='race')
```

```{python y new race}
y_new
```

# Exploratory Data Analysis
```{python exploratory visualization set up}
df_viz = X_new.copy()
df_viz['race'] = X_new['race'].replace({1.0: 'Caucasian', 0.0: 'African-American'})
df_viz['two_year_recid'] = y_new.replace({1:'Recidivated', 0: 'Survived'})
df_viz.index = df_viz.index.droplevel('race')

purple = '#9d0677'
green = '#30875c'
workshop_palette = [purple, green]
df_viz.head()
```

```{python exploratory recidivism bar plot}
# barplot of recividism
sns.countplot(x='two_year_recid', data=df_viz, palette=workshop_palette)
plt.title('Two Year Recidivism Rate')
```

```{python exploratory race bar plot}
# barplot of race
sns.countplot(y="race", data=df_viz, palette=["#30875c"])
plt.title('Race Distribution (White and Black individuals only)')
```

```{python exploratory age by race plot}
# age distribution by race
ax = sns.kdeplot(x="age", hue="race", data=df_viz);
kdeline = ax.lines[0]
mean_fem = df_viz.groupby('race').age.median()[0]
xs = kdeline.get_xdata()
ys = kdeline.get_ydata()
height_fem = np.interp(mean_fem, xs, ys)
ax.vlines(mean_fem, 0, height_fem, color=purple, ls=':')

mean_m = df_viz.groupby('race').age.median()[1]
height_m = np.interp(mean_m, xs, ys)
ax.vlines(mean_m, 0, height_m, color=green, ls=':')

print(df_viz.groupby('race').age.median())

plt.title('Distribution of Age by Race')
```

```{python exploratory recidivism by race plot}
# recidivism by race
by_sex = sns.countplot(x="race", hue="two_year_recid", data=df_viz, palette=workshop_palette)

plt.title('Two Year Recidivism by Race')
plt.show()
```

```{python exploratory recidivism by race table}
# table of recidivism by race
pd.crosstab(index = df_viz["race"], columns = df_viz["two_year_recid"])
```

# Baseline Metrics

```{python baseline train and test}
X_train, X_test, y_train, y_test = train_test_split(X_new, y_new, random_state=1234567)

# one-hoy encode the categorical features
data_preproc = make_column_transformer(
        (OneHotEncoder(sparse=False, handle_unknown='ignore'), X_train.dtypes == 'category'))

X_train = pd.DataFrame(data_preproc.fit_transform(X_train), index=X_train.index)
X_test = pd.DataFrame(data_preproc.transform(X_test), index=X_test.index)

# to save the information for the column names
pd.get_dummies(X_new).head()
```

```{python baseline set up, warning = FALSE, message = FALSE}
from aif360.sklearn.preprocessing import ReweighingMeta, Reweighing
from sklearn.metrics import accuracy_score, confusion_matrix, plot_roc_curve, recall_score, precision_score

lr = LogisticRegressionCV(solver='lbfgs')
reg = lr.fit(X_train, y_train)
y_pred = reg.predict(X_test)

acc_base = accuracy_score(y_test, y_pred)
print(f'[Baseline] The test accuracy of the algorithm is: {acc_base: .2%}')
```

```{python baseline confusion matrix}
##Look at 2nd column of matrix for precision 
##2nd row of matrix for recall 
cf_matrix = confusion_matrix(y_test, y_pred)
make_confusion_matrix(cf_matrix, "[Baseline]")
plt.show()
```

```{python baseline metrics per group}
metrics_per_group(y_test, y_pred)
plt.show()
```

In the Black population, the model correctly identifies 70% of people who recidivated. In the white population, the model correctly identifies 20% of people who recidivated.

### (Group fairness) Statistical Parity Difference

Computed as the difference of the rate of favorable outcomes (in this case, did not recidivate) received by the unprivileged group to the privileged group. It essentially equalizes the outcomes across the protected and non-protected groups.
The ideal value of this metric is 0. Fairness for this metric is between -0.1 and 0.1. A negative value means there is higher benefit for the privileged group (in this case, white defendants).

```{python group fairnesss statistical parity difference}
stat_par_diff = statistical_parity_difference(y_test, prot_attr='race', pos_label = 0)

print(f'[Baseline] The statistical parity difference is {stat_par_diff: .2}')

disp_impact_ratio = disparate_impact_ratio(y_test, y_pred, prot_attr='race', pos_label = 0)

print(f'[Baseline] The disparate impact ratio is {disp_impact_ratio: .2}')
# disparate impact ratio without y_pred attribute was 0.77

avg_odds_error = average_odds_error(y_test, y_pred, prot_attr='race', pos_label = 0)

print(f'[Baseline] The average odds error is {avg_odds_error: .2}')

avg_odds_diff = average_odds_difference(y_test, y_pred, prot_attr='race', pos_label = 0)

print(f'[Baseline] The average odds difference is {avg_odds_diff: .2}')
```

### Generalized entropy error

Generalized entropy index is proposed as a unified individual and group fairness measure. It measures the inequality in benefit allocation for individuals.
A value of 0 implies perfect fairness.
Fairness is indicated by lower scores, higher scores are problematic.

```{python group fairnesss generalized entropy error}
gen_entr_error = generalized_entropy_error(y_test, y_pred, alpha=1, pos_label = 0)

print(f'[Baseline] The generalized entropy error is {gen_entr_error: .2}')
```

### (Group fairness) Equal Opportunity Difference

This metric is computed as the difference of true positive rates between the unprivileged and the privileged groups. The true positive rate is the ratio of true positives to the total number of actual positives for a given group.
The ideal value is 0. A value of < 0 implies higher benefit for the privileged group and a value > 0 implies higher benefit for the unprivileged group.
Fairness for this metric is between -0.1 and 0.1

```{python group fairnesss equal opportunity difference}
eq_opp_diff = equal_opportunity_difference(y_test, y_pred, prot_attr='race', pos_label = 0)

print(f'[Baseline] The equal opportunity difference is {eq_opp_diff: .2}')

#bt_grp_gen_entr_error = between_group_generalized_entropy_error(y_test, y_pred, alpha=1, pos_label = 0)

#print(f'[Baseline] The between group generalized entropy error is {bt_grp_gen_entr_error: .2}')

#gen_entr_index = generalized_entropy_index(x_test, alpha=0, pos_label = 0)

#print(f'[Baseline] The generalized entropy index is {gen_entr_index: .2}')

#consistency_score = consistency_score(x_test, y_test, n_neighbors=5, pos_label = 0)

#print(f'[Baseline] The consistency score is {consistency_score: .2}')
```

```{python group fairness metrics plot}
plot_fair_metrics([0, 0, 0], '')
plt.show()
```

# Pre-Processing Approach
```{python pre-processing set up, warning = FALSE, message = FALSE}
lr = LogisticRegressionCV(solver='lbfgs')
rew = ReweighingMeta(estimator=lr, reweigher=Reweighing('race'))
rew.fit(X_train, y_train)
y_pred_REW = rew.predict(X_test)
```

```{python pre-processing accuracy score} 
acc_REW = accuracy_score(y_test, y_pred_REW)
print(f'[Reweighting] The test accuracy of the algorithm is: {acc_REW: .2%}')
```

```{python pre-processing confusion matrix}
cf_matrix = confusion_matrix(y_test, y_pred_REW)
make_confusion_matrix(cf_matrix, "[Reweighting]")
plt.show()
```

```{python pre-processing metrics per group}
metrics_per_group(y_test, y_pred_REW)
plt.show()
```

```{python pre-processing statistical parity difference}
stat_par_diff_RW = statistical_parity_difference(y_test, y_pred_REW, prot_attr='race', pos_label = 0)

print(f'[Reweighting] The statistical parity difference is {stat_par_diff_RW: .2}')
```

This is a large improvement over our baseline model, which was -0.14. It still implies a slight benefit for white individuals, but it is in the range of -0.1 and 0.1.

```{python pre-processing equal opportunity difference}
eq_opp_diff_RW = equal_opportunity_difference(y_test, y_pred_REW, prot_attr='race', pos_label = 0)

print(f'[Reweighting] The equal opportunity difference is {eq_opp_diff_RW: .2}')
```

This is also an improvement from the baseline, as it is now between -0.1 and 0.1. It also now implies a slight benefit for Black defendants.

```{python pre-processing generalized entropy error}
gen_entr_error_RW = generalized_entropy_error(y_test, y_pred_REW, alpha=1, pos_label = 0)

print(f'[Reweighting] The generalized entropy error is {gen_entr_error_RW: .2}')
```

This is only a very slight improvement from 0.24 to 0.23.

```{python pre-processing fairness metrics plot}
plot_fair_metrics([stat_par_diff_RW, eq_opp_diff_RW, gen_entr_error_RW], 'Reweighting')
plt.show()
```

# In-Processing

```{python in-processing set up}
df_train = X_train.copy()
df_train["two_year_recid"] = y_train
df_train.rename(columns={6:'race'}, inplace=True)
df_train = df_train.reset_index(drop=True)

df_test = X_test.copy()
df_test["two_year_recid"] = y_test
df_test = df_test.reset_index(drop=True)
df_test.rename(columns={6:'race'}, inplace=True)
```

```{python in-processing training}
train_BLD = BinaryLabelDataset(favorable_label='0',
                                unfavorable_label='1',
                                df=df_train,
                                label_names=['two_year_recid'],
                                protected_attribute_names=['race'])

test_BLD = BinaryLabelDataset(favorable_label='0',
                                unfavorable_label='1',
                                df=df_test,
                                label_names=['two_year_recid'],
                                protected_attribute_names=['race'])
```

```{python in-processing meta}
# meta = MetaFairClassifier(tau=0.8, sensitive_attr='race', type='fdr', seed=1234567)
# 
# meta.fit(train_BLD)
# y_pred_META = meta.predict(test_BLD)
# y_pred_META = y_pred_META.labels.flatten()
```

```{python in-processing accuracy score}
# acc_AD = accuracy_score(y_test, y_pred_META)
# print(f'[ Meta-Classification] The test accuracy of the algorithm is: {acc_AD: .2%}')
```

# Post-Processing

```{python post-processing calibrated equal odds set up, warning = FALSE, message = FALSE}
from aif360.sklearn.postprocessing import CalibratedEqualizedOdds, PostProcessingMeta

pp = CalibratedEqualizedOdds('race', cost_constraint='fnr', random_state=1234567)
ceo = PostProcessingMeta(estimator=lr, postprocessor=pp, random_state=1234567)
ceo.fit(X_train, y_train)
y_pred_CEO = ceo.predict(X_test)
y_proba_CEO = ceo.predict_proba(X_test)
```

```{python post-processing accuracy score}
acc_CEO = accuracy_score(y_test, y_pred_CEO)
print(f'[Calibrated Equalized Odds] The test accuracy of the algorithm is: {acc_CEO: .2%}')
```

```{python post-processing calibrated equal odds}
cf_matrix = confusion_matrix(y_test, y_pred_CEO)
make_confusion_matrix(cf_matrix, "[Calibrated Equalized Odds]")
plt.show()
```

```{python post-processing stat parity difference}
stat_par_diff_CEO = statistical_parity_difference(y_test, y_pred_CEO, prot_attr='race', pos_label = 0)

print(f'[Calibrated Equalized Odds] The statistical parity difference is {stat_par_diff_CEO: .2}')
```

```{python post-processing equal opportunity diff}
eq_opp_diff_CEO = equal_opportunity_difference(y_test, y_pred_CEO, prot_attr='race', pos_label = 0)

print(f'[Calibrated Equalized Odds] The equal opportunity difference is {eq_opp_diff_CEO: .2}')
```

```{python post-processing generalized entropy error}
gen_entr_error_CEO = generalized_entropy_error(y_test, y_pred_CEO, alpha=1, pos_label = 0)

print(f'[Calibrated Equalized Odds] The generalized entropy error is {gen_entr_error_CEO: .2}')
```

```{python post-processing fairness metrics plot}
plot_fair_metrics([stat_par_diff_CEO, eq_opp_diff_CEO, gen_entr_error_CEO], 'Calibrated Equalized Odds')
plt.show()
```