### 2022-04-13 final_metrics_analysis.Rmd

---
title: "Analyzing Fairness: Race, Recidivism, and the COMPAS Algorithm"
author: "Elisabeth Nesmith, Eunice Kim, Emma Kornberg, Amrita Acharya, Dianne Caravela"
date: "3/8/2022"
output: html_document
---

# Set Up

```{r py setup, include=FALSE}
library(reticulate)
```

```{python import, include= FALSE}
import matplotlib.pyplot as plt
import matplotlib.patches as patches
import numpy as np
import pandas as pd
import seaborn as sns

import matplotlib.pyplot as plt
import matplotlib.patches as mpatches

from sklearn.compose import make_column_transformer
from sklearn.linear_model import LogisticRegression
from sklearn.linear_model import LogisticRegressionCV

from sklearn.model_selection import GridSearchCV, train_test_split
from sklearn.preprocessing import OneHotEncoder

from aif360.datasets import CompasDataset
from aif360.datasets import BinaryLabelDataset
from aif360.sklearn.datasets import fetch_compas
from aif360.algorithms.inprocessing import MetaFairClassifier

from aif360.sklearn.preprocessing import ReweighingMeta
from aif360.sklearn.inprocessing import AdversarialDebiasing
from aif360.sklearn.postprocessing import CalibratedEqualizedOdds, PostProcessingMeta

from aif360.sklearn.metrics import disparate_impact_ratio, average_odds_difference, statistical_parity_difference, equal_opportunity_difference

# from aif360.algorithms.preprocessing import DisparateImpactRemover
# from BlackBoxAuditing.repairers.GeneralRepairer import Repairer

from IPython.display import Markdown, display
```

# Data

```{python fetch compas}
## load in data
## In machine learning tasks, specifically with supervised learning, you have features and labels. 
## The features are the descriptive attributes (they are defined as X), and the label (y) is what you're attempting to predict or forecast

X, y = fetch_compas()
print(f'There are {X.shape[0]} entries and {X.shape[1]} features')
X.head()
```

## Filter/configure race

```{python filter race}
## because our analysis is mainly focusing on how the algorithm treats white and Black people differently, we are
## dropping the rows of data where race != Caucasian or African American
X_new = X[(X.race == "Caucasian") | (X.race == "African-American")]
print(f'There are {X_new.shape[0]} entries and {X_new.shape[1]} features')
X_new.head()
```

```{python drop other races}
## drop unused race categories
# list of categories to be removed
X_new["race"] = X_new["race"].cat.remove_unused_categories()
```

```{python configure races}
y_new = y[(y.index.get_level_values(2) == "Caucasian") | (y.index.get_level_values(2) == "African-American")]
y_new.head()
```

```{python initial confusion matrix}
# Function for visualising the confusion matrix and other statistics
# https://github.com/DTrimarchi10/confusion_matrix/blob/master/cf_matrix.py

def make_confusion_matrix(cf_matrix, model):
  group_names = ["True Negative","False Positive","False Negative","True Positive"]
  group_counts = ["{0:0.0f}".format(value) for value in
                  cf_matrix.flatten()]
  group_percentages = ["{0:.2%}".format(value) for value in
                      cf_matrix.flatten()/np.sum(cf_matrix)]

  group_labels = ["{}\n".format(value) for value in group_names]
  group_counts = ["{0:0.0f}\n".format(value) for value in cf_matrix.flatten()]
  group_percentages = ["{0:.2%}".format(value) for value in cf_matrix.flatten()/np.sum(cf_matrix)]

  box_labels = [f"{v1}{v2}{v3}".strip() for v1, v2, v3 in zip(group_labels,group_counts,group_percentages)]
  box_labels = np.asarray(box_labels).reshape(cf_matrix.shape[0],cf_matrix.shape[1])

  # add more statistics
  accuracy  = np.trace(cf_matrix) / float(np.sum(cf_matrix))
  precision = cf_matrix[1,1] / sum(cf_matrix[:,1])
  recall    = cf_matrix[1,1] / sum(cf_matrix[1,:])
  f1_score  = 2*precision*recall / (precision + recall)
  stats_text = "\n\nAccuracy={:0.3f}\nPrecision={:0.3f}\nRecall={:0.3f}\nF1 Score={:0.3f}".format(
      accuracy,precision,recall,f1_score)

  categories=["Survived", "Recidivated"]
  sns.heatmap(cf_matrix,annot=box_labels,fmt="",cmap='Purples',xticklabels=categories,yticklabels=categories)

  plt.ylabel('True label')
  plt.xlabel('Predicted label' + stats_text)
  plt.title(f"Confusion matrix and statistics for the {model} model");

## defining function for displaying metrics of training and test data by sex 
def metrics_per_group(y_test, y_pred):
	# y true per group
	y_test_white = y_test.loc[y_test.index.get_level_values(2) == 1]
	y_test_black = y_test.loc[y_test.index.get_level_values(2) == 0]

	# y_pred per group
	y_pred_white = y_pred[y_test.index.get_level_values(2) == 1]
	y_pred_black = y_pred[y_test.index.get_level_values(2) == 0]

	# metrics
	scores = []
	scores.append(accuracy_score(y_test, y_pred))
	scores.append(recall_score(y_test, y_pred))
	scores.append(precision_score(y_test, y_pred))

	scores.append(accuracy_score(y_test_black, y_pred_black))
	scores.append(recall_score(y_test_black, y_pred_black))
	scores.append(precision_score(y_test_black, y_pred_black))

	scores.append(accuracy_score(y_test_white, y_pred_white))
	scores.append(recall_score(y_test_white, y_pred_white))
	scores.append(precision_score(y_test_white, y_pred_white))

	attribute = ["all"]*3 + ["black"] *3 + ["white"] *3
	metric = ["accuracy", "recall", "precision"] * 3

	# dictionary of lists 
	dict = {'race': attribute, 'metrics': metric, 'score': scores}

	df = pd.DataFrame(dict)

	sns.barplot(x = "metrics", y = "score", hue = "race", data = df, palette = ['#dfcd1a', '#9d0677', '#236c48'])
	plt.title("Performance metrics by groups")

def plot_fair_metrics(fair_metrics_mitigated, model): 
  cols = ['statistical_parity_difference','equal_opportunity_difference','average_odds_difference','disparate_impact_ratio']
  obj_fairness = [[0,0,0,1]]

  # row for objectives    
  fair_metrics = pd.DataFrame(data=obj_fairness, index=['objective'], columns=cols)

  # row for baseline model
  fair_metrics.loc['Baseline Model'] = [stat_par_diff, eq_opp_diff, avg_odds_diff, disp_impact_ratio]

  # row for mitigated bias
  fair_metrics.loc[model] = fair_metrics_mitigated

  metrics_len = len(cols)

  fig, ax = plt.subplots(figsize=(20,4), ncols=metrics_len, nrows=1)

  plt.subplots_adjust(
      left    =  0.125, 
      bottom  =  0.1, 
      right   =  0.9, 
      top     =  0.9, 
      wspace  =  .5, 
      hspace  =  1.1
  )

  y_title_margin = 1.2

  plt.suptitle("Fairness metrics", y = 1.09, fontsize=20)
  sns.set(style="dark")

  cols = fair_metrics.columns.values
  obj = fair_metrics.loc['objective']
  size_rect = [0.2,0.2,0.2,0.4]
  rect = [-0.1,-0.1,-0.1,0.8]
  bottom = [-1,-1,-1,0]
  top = [1,1,1,2]
  bound = [[-0.1,0.1],[-0.1,0.1],[-0.1,0.1],[0.8,1.25]]

  for i in range(0,metrics_len):
      plt.subplot(1, metrics_len, i+1)
      ax = sns.barplot(x=fair_metrics.index[1:len(fair_metrics)], y=fair_metrics.iloc[1:len(fair_metrics)][cols[i]])

      for j in range(0,len(fair_metrics)-1):
          a, val = ax.patches[j], fair_metrics.iloc[j+1][cols[i]]
          marg = -0.2 if val < 0 else 0.1
          ax.text(a.get_x()+a.get_width()/5, a.get_y()+a.get_height()+marg, round(val, 3), fontsize=15,color='black')

      plt.ylim(bottom[i], top[i])
      plt.setp(ax.patches, linewidth=0)
      ax.add_patch(patches.Rectangle((-5,rect[i]), 10, size_rect[i], alpha=0.3, facecolor="green", linewidth=1, linestyle='solid'))
      plt.axhline(obj[i], color='black', alpha=0.3)
      plt.title(cols[i])
      ax.set_ylabel('')    
      ax.set_xlabel('')
```

```{python new race index}
X_new.index = pd.MultiIndex.from_arrays(X_new.index.codes, names=X_new.index.names)
y_new.index = pd.MultiIndex.from_arrays(y_new.index.codes, names=y_new.index.names)
# 0 is African American, 2 is Caucasian
```

```{python re-index race}
# set caucasian equal to 1 instead of 2
X_new = X_new.rename(index={2: 1}, level='race')
X_new
```

```{python target class}
# set target class to 0/1 
y_new = pd.Series(y_new.factorize(sort=True)[0], index=y_new.index)
# set caucasian equal to 1 instead of 2
y_new = y_new.rename(index={2: 1}, level='race')
```

```{python y new race}
y_new
```

# Exploratory Data Analysis
```{python exploratory visualization set up}
df_viz = X_new.copy()
df_viz['race'] = X_new['race'].replace({1.0: 'Caucasian', 0.0: 'African-American'})
df_viz['two_year_recid'] = y_new.replace({1:'Recidivated', 0: 'Survived'})
df_viz.index = df_viz.index.droplevel('race')

purple = '#9d0677'
green = '#30875c'
workshop_palette = [purple, green]
df_viz.head()
```

```{python exploratory recidivism bar plot}
# barplot of recividism
sns.countplot(x='two_year_recid', data=df_viz, palette=workshop_palette)
plt.title('Two Year Recidivism Rate')
```

```{python exploratory race bar plot}
# barplot of race
sns.countplot(y="race", data=df_viz, palette=["#30875c"])
plt.title('Race Distribution (White and Black individuals only)')
```

```{python exploratory age by race plot}
# age distribution by race
ax = sns.kdeplot(x="age", hue="race", data=df_viz);
kdeline = ax.lines[0]
mean_fem = df_viz.groupby('race').age.median()[0]
xs = kdeline.get_xdata()
ys = kdeline.get_ydata()
height_fem = np.interp(mean_fem, xs, ys)
ax.vlines(mean_fem, 0, height_fem, color=purple, ls=':')

mean_m = df_viz.groupby('race').age.median()[1]
height_m = np.interp(mean_m, xs, ys)
ax.vlines(mean_m, 0, height_m, color=green, ls=':')

print(df_viz.groupby('race').age.median())

plt.title('Distribution of Age by Race')
```

```{python exploratory recidivism by race plot}
# recidivism by race
by_sex = sns.countplot(x="race", hue="two_year_recid", data=df_viz, palette=workshop_palette)

plt.title('Two Year Recidivism by Race')
plt.show()
```

```{python exploratory recidivism by race table}
# table of recidivism by race
pd.crosstab(index = df_viz["race"], columns = df_viz["two_year_recid"])
```

# Baseline Metrics

```{python baseline train and test}
X_train, X_test, y_train, y_test = train_test_split(X_new, y_new, random_state=1234567)

# one-hoy encode the categorical features
data_preproc = make_column_transformer(
        (OneHotEncoder(sparse=False, handle_unknown='ignore'), X_train.dtypes == 'category'))

X_train = pd.DataFrame(data_preproc.fit_transform(X_train), index=X_train.index)
X_test = pd.DataFrame(data_preproc.transform(X_test), index=X_test.index)

# to save the information for the column names
pd.get_dummies(X_new).head()
```

```{python baseline set up, include = FALSE}
from aif360.sklearn.preprocessing import ReweighingMeta, Reweighing
from sklearn.metrics import accuracy_score, confusion_matrix, plot_roc_curve, recall_score, precision_score

lr = LogisticRegressionCV(solver='lbfgs')
reg = lr.fit(X_train, y_train)
y_pred = reg.predict(X_test)

acc_base = accuracy_score(y_test, y_pred)
print(f'[Baseline] The test accuracy of the algorithm is: {acc_base: .2%}')
```

```{python baseline confusion matrix}
##Look at 2nd column of matrix for precision 
##2nd row of matrix for recall 
cf_matrix = confusion_matrix(y_test, y_pred)
make_confusion_matrix(cf_matrix, "[Baseline]")
plt.show()
```

```{python baseline metrics per group}
metrics_per_group(y_test, y_pred)
plt.show()
```

In the Black population, the model correctly identifies 70% of people who recidivated. In the white population, the model correctly identifies 20% of people who recidivated.

## Group Fairness Metrics

With group fairness metrics, different groups should receive similar treatments or outcomes.  In this case of recidivism and race, this means that Black defendants should have similar rates of predicted recidivism as white defendants.

### Statistical Parity Difference

Statistical Parity difference is computed as the difference in the rate of favorable outcomes (in this case, did not recidivate) received by the unprivileged group to the privileged group. It essentially equalizes the outcomes across the privileged and non-privileged groups.
The ideal value of this metric is 0. Fairness for this metric is between -0.1 and 0.1. A negative value means there is higher benefit for the privileged group (in this case, white defendants).

$P(\hat{Y}=1|D=Unprivileged) - P(\hat{Y}=1|D=Privileged)$

```{python group fairnesss statistical parity difference}
stat_par_diff = statistical_parity_difference(y_test, prot_attr='race', pos_label = 0)

print(f'[Baseline] The statistical parity difference is {stat_par_diff: .2}')
```

Because the statistical parity difference is negative and not within -0.1 and 0.1, the algorithm unfairly benefits white defendants.

### Disparate Impact Ratio

This metric is the ratio of how often the favorable outcome occurs in one group versus the other. In the case of recidivism, this is the ratio of how many white defendants are predicted to not recidivate compared to how many black defendants are predicted to not recidivate.

A value of 1 means that the ratio is exactly 1:1. Less than 1 means the privileged group (white defendants) benefits, while a value greater than 1 means the unprivileged group (Black defendants) benefits. According to AI Fairness 360, a ratio between .8 to 1.25 is considered fair.

The disparate impact ratio is calculated with the formula:

$\frac{P(\hat{Y}=1|D=Unprivileged)}{P(\hat{Y}=1|D=Privileged)}$

where Y is the favorable outcome of a defendant's predicted two year recidivism and D is the race of the defendant. In our case, the favorable outcome is survived, or not recidivated, which is Y = 0. In terms of race, 0 = Black and 1 = White. So to calculate the disparate impact ratio on this data, the formula is:

$\frac{P(\hat{Y}=0|D=0)}{P(\hat{Y}=0|D=1)}$

Say we have a data test set of 200 defendants, 125 Black and 75 white. Of the 125 Black defendants, 44% (55) were predicted to not recidivate. Of the 75 white defendants, 66% (50) were predicted to not recidivate.

The disparate impact ratio would therefore be

$\frac{0.44}{0.66} = 0.667$

Because this value is less than 1, it benefits white defendants. Additionally, because it is below 0.8, it is considered "unfair."

```{python group fairnesss disparate impact ratio}
disp_impact_ratio = disparate_impact_ratio(y_test, y_pred, prot_attr='race', priv_group = 1, pos_label = 0)

print(f'[Baseline] The disparate impact ratio is {disp_impact_ratio: .2}')
```

Since the actual disparate impact ratio for our test data is 0.47 and below 0.8, the algorithm unfairly benefits white defendants i.e. disproportionately predicts them to not recidivate.

### Average Odds Difference

This metric returns the average difference in false positive rate and true positive rate for the privileged and unprivileged groups. A value of 0 is considered fair, and a value below 0 implies benefit for the privileged group. Equality of odds is achieved in the case of recidivism when the proportion of people who were predicted to recidivate and did recidivate is equal (true positive rate) for both black and white defendants AND the proportion of people who were predicted to recidivate and did not recidivate (false positive rate) is equal for both black and white defendants.

$\frac{(FPR_{D = unprivileged} - FPR_{D = privileged}) + (TPR_{D = unprivileged} - TPR_{D = privileged})}{2}$

```{python group fairnesss average odds difference}
avg_odds_diff = average_odds_difference(y_test, y_pred, prot_attr='race', pos_label = 0)

print(f'[Baseline] The average odds difference is {avg_odds_diff: .2}')
```

Since the average odds difference is below 0 by a substantial amount, white defendants are unfairly benefitted by the algorithm.

### Equal Opportunity Difference

This metric is computed as the difference of true positive rates between the unprivileged and the privileged groups. The true positive rate is the ratio of true positives to the total number of actual positives for a given group.
The ideal value is 0. A value of < 0 implies higher benefit for the privileged group and a value > 0 implies higher benefit for the unprivileged group.
Fairness for this metric is between -0.1 and 0.1.

$TPR_{D = Unprivileged} - TPR_{D = Privileged}$

```{python group fairnesss equal opportunity difference}
eq_opp_diff = equal_opportunity_difference(y_test, y_pred, prot_attr='race', pos_label = 0)

print(f'[Baseline] The equal opportunity difference is {eq_opp_diff: .2}')
```

Since the equal opportunity difference is below 0 and beyond -0.1, the algorithm unfairly benefits white defendants.

### Plot Fairness Metrics (Baseline)

```{python group fairness metrics plot}
plot_fair_metrics([0, 0, 0, 0], '')
plt.show()
```

# Pre-Processing Approaches

## Reweighing

```{python pre-processing set up, include = FALSE}
lr = LogisticRegressionCV(solver='lbfgs')
rew = ReweighingMeta(estimator=lr, reweigher=Reweighing('race'))
rew.fit(X_train, y_train)
y_pred_REW = rew.predict(X_test)
```

```{python pre-processing accuracy score} 
acc_REW = accuracy_score(y_test, y_pred_REW)
print(f'[Reweighting] The test accuracy of the algorithm is: {acc_REW: .2%}')
```

```{python pre-processing confusion matrix}
cf_matrix = confusion_matrix(y_test, y_pred_REW)
make_confusion_matrix(cf_matrix, "[Reweighting]")
plt.show()
```

```{python pre-processing metrics per group}
metrics_per_group(y_test, y_pred_REW)
plt.show()
```

### Statistical Parity Difference

```{python pre-processing statistical parity difference}
stat_par_diff_RW = statistical_parity_difference(y_test, y_pred_REW, prot_attr='race', pos_label = 0)

print(f'[Reweighting] The statistical parity difference is {stat_par_diff_RW: .2}')
```

This is a large improvement over our baseline model, which was -0.14. It still implies a slight benefit for white individuals, but it is in the range of -0.1 and 0.1.

### Equal Opportunity Difference

```{python pre-processing equal opportunity difference}
eq_opp_diff_RW = equal_opportunity_difference(y_test, y_pred_REW, prot_attr='race', pos_label = 0)

print(f'[Reweighting] The equal opportunity difference is {eq_opp_diff_RW: .2}')
```

This is also an improvement from the baseline, as it is now between -0.1 and 0.1. It also now implies a slight benefit for Black defendants.

### Average Odds Difference

```{python pre-processing average odds difference}
avg_odds_diff_RW = average_odds_difference(y_test, y_pred_REW, prot_attr='race', pos_label = 0)

print(f'[Reweighting] The average odds difference is {avg_odds_diff_RW: .2}')
```

This is also a vast improvement from -0.44 to 0.014. Since it is above 1 and below 0.1, it implies a slight benefit for Black defendants, but not an unfair benefit.

### Disparate Impact Ratio
```{python pre_processing disparate impact ratio}
disp_impact_ratio_RW = disparate_impact_ratio(y_test, y_pred_REW, prot_attr='race', priv_group = 1, pos_label = 0)

print(f'[Reweighting] The disparate impact ratio is {disp_impact_ratio_RW: .2}')
```

The reweighting approach also improves the disparate impact ratio, bringing the value almost to 1, which conveys absolute fairness.

### Plot Fairness Metrics (Reweighting)

```{python pre-processing fairness metrics plot}
plot_fair_metrics([stat_par_diff_RW, eq_opp_diff_RW, avg_odds_diff_RW, disp_impact_ratio_RW], 'Reweighting')
plt.show()
```

```{python pre-processing disparate impact remover}
#lr = LogisticRegressionCV(solver='lbfgs')
#rew = ReweighingMeta(estimator=lr, reweigher=Reweighing('race'))
#rew.fit(X_train, y_train)
#y_pred_REW = rew.predict(X_test)

# diremover = DisparateImpactRemover(repair_level=1.0, sensitive_attribute="race")
#diremover.fit(X_train, y_train)
#y_pred_diremover = diremover.predict(X_test)

# diremover_rp = diremover.fit_transform(X_test)
# AttributeError: 'DataFrame' object has no attribute 'features' (probem with .fit_transform function)
# .fit_transform() angry because of X_test... 
```

# In-Processing

```{python in-processing set up}
df_train = X_train.copy()
df_train["two_year_recid"] = y_train
df_train.rename(columns={6:'race'}, inplace=True)
df_train = df_train.reset_index(drop=True)

df_test = X_test.copy()
df_test["two_year_recid"] = y_test
df_test = df_test.reset_index(drop=True)
df_test.rename(columns={6:'race'}, inplace=True)
```

```{python in-processing training}
train_BLD = BinaryLabelDataset(favorable_label='0',
                                unfavorable_label='1',
                                df=df_train,
                                label_names=['two_year_recid'],
                                protected_attribute_names=['race'])

test_BLD = BinaryLabelDataset(favorable_label='0',
                                unfavorable_label='1',
                                df=df_test,
                                label_names=['two_year_recid'],
                                protected_attribute_names=['race'])
```

```{python in-processing meta}
# meta = MetaFairClassifier(tau=0.8, sensitive_attr='race', type='fdr', seed=1234567)
# 
# meta.fit(train_BLD)
# y_pred_META = meta.predict(test_BLD)
# y_pred_META = y_pred_META.labels.flatten()
```

```{python in-processing accuracy score}
# acc_AD = accuracy_score(y_test, y_pred_META)
# print(f'[ Meta-Classification] The test accuracy of the algorithm is: {acc_AD: .2%}')
```

# Post-Processing

```{python post-processing calibrated equal odds set up, include = FALSE}
from aif360.sklearn.postprocessing import CalibratedEqualizedOdds, PostProcessingMeta

pp = CalibratedEqualizedOdds('race', cost_constraint='fnr', random_state=1234567)
ceo = PostProcessingMeta(estimator=lr, postprocessor=pp, random_state=1234567)
ceo.fit(X_train, y_train)
y_pred_CEO = ceo.predict(X_test)
y_proba_CEO = ceo.predict_proba(X_test)
```

```{python post-processing accuracy score}
acc_CEO = accuracy_score(y_test, y_pred_CEO)
print(f'[Calibrated Equalized Odds] The test accuracy of the algorithm is: {acc_CEO: .2%}')
```

```{python post-processing calibrated equal odds}
cf_matrix = confusion_matrix(y_test, y_pred_CEO)
make_confusion_matrix(cf_matrix, "[Calibrated Equalized Odds]")
plt.show()
```

### Statistical Parity Difference

```{python post-processing stat parity difference}
stat_par_diff_CEO = statistical_parity_difference(y_test, y_pred_CEO, prot_attr='race', pos_label = 0)

print(f'[Calibrated Equalized Odds] The statistical parity difference is {stat_par_diff_CEO: .2}')
```

### Equal Opportunity Difference

```{python post-processing equal opportunity diff}
eq_opp_diff_CEO = equal_opportunity_difference(y_test, y_pred_CEO, prot_attr='race', pos_label = 0)

print(f'[Calibrated Equalized Odds] The equal opportunity difference is {eq_opp_diff_CEO: .2}')
```

### Average Odds Difference

```{python post-processing average odds difference}
avg_odds_diff_CEO = average_odds_difference(y_test, y_pred_CEO, prot_attr='race', pos_label = 0)

print(f'[Calibrated Equalized Odds] The average odds difference is {avg_odds_diff_CEO: .2}')
```

### Disparate Impact Ratio

```{python post_processing disparate impact ratio}
disp_impact_ratio_CEO = disparate_impact_ratio(y_test, y_pred_CEO, prot_attr='race', priv_group = 1, pos_label = 0)

print(f'[Calibrated Equalized Odds] The disparate impact ratio is {disp_impact_ratio_CEO: .2}')
```

### Plot Fairness Metrics (Calibrated Equalized Odds)

```{python post-processing fairness metrics plot}
plot_fair_metrics([stat_par_diff_CEO, eq_opp_diff_CEO, avg_odds_diff_CEO, disp_impact_ratio_CEO], 'Calibrated Equalized Odds')
plt.show()
```